---
title: "Project 2"
author: "Maria Camila Restrepo"
date: "2024-09-20"
output: html_document
---

```{r setup}
library(tidyverse)
library(here)
```

# Part 1: Fun with functions

### Part 1A: Exponential transformation

The exponential of a number can be written as an infinite series expansion of the form $$
\exp(x) = 1 + x + \frac{x^2}{2!} + \frac{x^3}{3!} + \cdots
$$ Of course, we cannot compute an infinite series by the end of this term and so we must truncate it at a certain point in the series. The truncated sum of terms represents an approximation to the true exponential, but the approximation may be usable.

Write a function that computes the exponential of a number using the truncated series expansion. The function should take two arguments:

-   `x`: the number to be exponentiated

-   `k`: the number of terms to be used in the series expansion beyond the constant 1. The value of `k` is always $\geq 1$.

For example, if $k = 1$, then the `Exp` function should return the number $1 + x$. If $k = 2$, then you should return the number $1 + x + x^2/2!$.

Include at least one example of output using your function.

```{r}

trunc_exp <- function(x, k) {
  
  result <- 1 # Initialize the result with the constant 1 (the 0th term)
  
  # Loop through terms from 1 to k to compute the series sum
  for (i in 1:k) {
        result <- result + (x^i) / factorial(i)
  }
  
  return(result)
}
```

Result:

```{r}
trunc_exp(2, 50) 
```

### Part 1B: Sample mean and sample standard deviation

Next, write two functions called `sample_mean()` and `sample_sd()` that takes as input a vector of data of length $N$ and calculates the sample average and sample standard deviation for the set of $N$ observations.

```{r}
sample_mean <- function(x) {
  return(sum(x) / length(x))
}

sample_sd <- function(x) {
  mean <- sample_mean(x)
  accumulator <- 0 #need to start with the 0th term for the sum
  for (i in x) { #i is the index (each number position inside the vector)
    accumulator <- accumulator + (i - mean)^2
  }
  result <- sqrt ( accumulator / (length(x) - 1) )
  return(result)
}
```

Result:

```{r}
vec <- c(1:10)
sample_mean(vec)
sample_sd(vec)
sd(vec) #checking the result
```

# Part 2: Wrangling data  

```{r, message=FALSE}
library(here)
if (!file.exists(here("data", "tuesdata_rainfall.RDS"))) {
    tuesdata <- tidytuesdayR::tt_load("2020-01-07")
    rainfall <- tuesdata$rainfall
    temperature <- tuesdata$temperature

    # save the files to RDS objects
    saveRDS(tuesdata$rainfall, file = here("data", "tuesdata_rainfall.RDS"))
    saveRDS(tuesdata$temperature, file = here("data", "tuesdata_temperature.RDS"))
}
```
Let's load the datasets

```{r, eval=TRUE, message=FALSE}
rainfall <- readRDS(here("data", "tuesdata_rainfall.RDS"))
temperature <- readRDS(here("data", "tuesdata_temperature.RDS"))
```

Now we can look at the data with `glimpse()`

```{r,message=FALSE}

glimpse(rainfall)
glimpse(temperature)
```

this dataset contains temperature and rainfall data from Australia.

Here is a data dictionary for what all the column names mean:

-   <https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-07/readme.md#data-dictionary>

### Tasks

Using the `rainfall` and `temperature` data, perform the following steps and create a new data frame called `df`:

1.  Start with `rainfall` dataset and drop any rows with NAs.
```{r}
nona_rainfall <- rainfall %>%
  drop_na()
#now its 64,476 obs
```

2.  Create a new column titled `date` that combines the columns `year`, `month`, `day` into one column separated by "-". (e.g. "2020-01-01"). This column should not be a character, but should be recognized as a date. (**Hint**: check out the `ymd()` function in `lubridate` R package). You will also want to add a column that just keeps the `year`.

```{r}
library(lubridate)
library(dplyr)
library(stringr)

rainfall1 <- nona_rainfall %>%
  mutate(date_str = str_c(year, month, day, sep = "-")) %>%
  mutate(date = ymd(date_str)) %>%
  select(-date_str, -month, -day)

glimpse(rainfall1)
```


3.  Using the `city_name` column, convert the city names (character strings) to all upper case.
4.  Join this wrangled rainfall dataset with the `temperature` dataset such that it includes only observations that are in both data frames. (**Hint**: there are two keys that you will need to join the two datasets together). (**Hint**: If all has gone well thus far, you should have a dataset with 83,964 rows and 13 columns).

::: callout-note
-   You may need to use functions outside these packages to obtain this result, in particular you may find the functions `drop_na()` from `tidyr` and `str_to_upper()` function from `stringr` useful.
:::

```{r}
# Add your solution here
```
