---
title: "Project 1"
author: "Maria Camila Restrepo"
date: "2024-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(usethis)
library("tidyverse")

library(remotes)
  if (!require("tidytuesdayR", quietly = TRUE)) {
    remotes::install_github("dslc-io/tidytuesdayR")
  }

library("tidytuesdayR")  
library("here")

```

## Setup
```{r}

# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}

# saves data only once (not each time you knit a R Markdown)
if (!file.exists(here("data", "chocolate.RDS"))) {
    url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv"
    chocolate <- readr::read_csv(url_csv)

    # save the file to RDS objects
    saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}

chocolate <- readRDS(here("data", "chocolate.RDS"))

as_tibble(chocolate)
glimpse(chocolate)
```
Here is a data dictionary for what all the column names mean:

https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-01-18/readme.md#data-dictionary

Here is a data dictionary for what all the column names mean:

https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-01-18/readme.md#data-dictionary

## Part 1: Explore Data

In this part, use functions from dplyr and ggplot2 to answer the following questions.

1.  Make a histogram of the `rating` scores to visualize the overall distribution of scores. Change the number of bins from the default to 10, 15, 20, and 25. Pick on the one that you think looks the best. Explain what the difference is when you change the number of bins and explain why you picked the one you did.

```{r}
#install.packages("esquisse")
#library(esquisse)
#esquisser(chocolate)

#ggplot(chocolate) +
#  aes(x = rating) +
#  geom_histogram(bins = 30L, fill = "#112446") +
#  theme_minimal()

ggplot(chocolate) +
 aes(x = rating) +
 geom_histogram(bins = 10)

ggplot(chocolate) +
 aes(x = rating) +
 geom_histogram(bins = 15)

ggplot(chocolate) +
 aes(x = rating) +
 geom_histogram(bins = 20)

ggplot(chocolate) +
 aes(x = rating) +
 geom_histogram(bins = 25)

```
  
Specifying the number of bins in the geom_histogram function controls the number of bins in the histogram. I would choose the histograms with 15 bins because the values are not continuous, making it unnecessary to divide into more bins.

```{r}
ggplot(chocolate) +
 aes(x = rating) +
 geom_histogram(bins = 15)
```


2.  Consider the countries where the beans originated from. How many reviews come from each country of bean origin?

```{r}
chocolate %>%
  group_by(country_of_bean_origin) %>%
  summarize(count = n())
```
3.  What is average `rating` scores from reviews of chocolate bars that have Ecuador as `country_of_bean_origin` in this dataset? For this same set of reviews, also calculate (1) the total number of reviews and (2) the standard deviation of the `rating` scores. Your answer should be a new data frame with these three summary statistics in three columns. Label the name of these columns `mean`, `sd`, and `total`.

```{r}
chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(country_of_bean_origin == "Ecuador") %>%
  summarize(average_raiting = mean(rating),
            total_reviews = n(),
            sd_rating = sd(rating))
```

4.  Which company (name) makes the best chocolate (or has the highest ratings on average) with beans from Ecuador?

```{r}
chocolate %>%
  filter(country_of_bean_origin == "Ecuador") %>%
  group_by(company_manufacturer) %>%
  summarize(average_rating = mean(rating)) %>%
  arrange(desc(average_rating))
```

The companies that make the best chocolate with beans from Ecuador are Amano, Benoit Nihant, Beschle (Felchlin), Durci and The Smooth Chocolator.

5.  Calculate the average rating across all country of origins for beans. Which top 3 countries (for bean origin) have the highest ratings on average?

```{r}
chocolate %>%
  group_by(country_of_bean_origin) %>%
  summarize(average_rating = mean(rating)) %>%
  arrange(desc(average_rating))
```

The top three countries with the highest average rating are Tobago, China and Sao Tome & Principe.

6.  Following up on the previous problem, now remove any countries of bean origins that have less than 10 chocolate bar reviews. Now, which top 3 countries have the highest ratings on average?

```{r}
chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 10) %>%
  summarize(average_rating = mean(rating)) %>%
  arrange(desc(average_rating))
```

Now the top three countries are Solomon Islands, Congo and Cuba.


7.  For this last part, let's explore the relationship between percent chocolate and ratings.

Use the functions in `dplyr`, `tidyr`, and `lubridate` to perform the following steps to the `chocolate` dataset:

1.  Identify the countries of bean origin with at least 50 reviews. Remove reviews from countries are not in this list.
2.  Using the variable describing the chocolate percentage for each review, create a new column that groups chocolate percentages into one of four groups: (i) \<60%, (ii) \>=60 to \<70%, (iii) \>=70 to \<90%, and (iii) \>=90% (**Hint** check out the `substr()` function in base R and the `case_when()` function from `dplyr` -- see example below).
3.  Using the new column described in #2, re-order the factor levels (if needed) to be starting with the smallest percentage group and increasing to the largest percentage group (**Hint** check out the `fct_relevel()` function from `forcats`).


```{r}
library(forcats)

percentage <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 50) %>%
  mutate(chocolate_percentage = case_when(
    cocoa_percent < 60 ~ "<60%",
    cocoa_percent >= 60 & cocoa_percent <70 ~ ">=60 to <70%", 
    cocoa_percent >= 70 & cocoa_percent <90 ~ ">=70 to <90%",
    cocoa_percent >= 90 ~ ">=90%"
    )) %>%
  mutate(chocolate_percentage = as.factor(chocolate_percentage))

str(percentage) #checking if chocolate_percentage was converted into a factor. However, the factors are in the wrong order

percentage$chocolate_percentage <- factor(percentage$chocolate_percentage, levels= c("<60%", ">=60 to <70%", ">=70 to <90%", ">90%")) #I used factor instead of fct_relevel

levels(percentage$chocolate_percentage)
str(percentage$chocolate_percentage) #the factors are ordered correctly now


```

4.  For each country, make a set of four side-by-side boxplots plotting the groups on the x-axis and the ratings on the y-axis. These plots should be faceted by country.

```{r}

chocolate %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    num_reviews = n(),
    unique_countries = n_distinct(country_of_bean_origin)
  )#62 countries

percentage %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    num_reviews = n(),
    unique_countries = n_distinct(country_of_bean_origin)
  )#16 countries left

#Checking which countries of origin are left in the dataset after filtering out those with less than 50 reviews.

ggplot(percentage) +  
  aes(x = chocolate_percentage, y = rating) +
  geom_boxplot() #without faceting per country

ggplot(percentage) +  
  aes(x = chocolate_percentage, y = rating) +
  geom_boxplot() +
  facet_wrap(~ country_of_bean_origin)

percentage %>%
  group_by(chocolate_percentage) %>%
  summarize(average_rating = mean(rating))

```

On average, which category of chocolate percentage is most highly rated? Do these countries mostly agree or are there disagreements?

The most highly rated category is >=60 to <70% cacao percentage, but the average rating is very similar to chocolate with >=70 to <90% cacao percentage. In most countries, it is noticable that the category with the lowest rating is <60% cacao percentage, in comparison to the other two. The magnitude of the difference between the second and third categories varies between countries. 

# Part 2: Join two datasets together

### Tasks

1.  Use this dataset it to create a new column called `continent` in our `chocolate` dataset that contains the continent name for each review where the country of bean origin is.

```{r}
#install.packages("gapminder")
library(gapminder)
gapminder
str(gapminder)


gm_country <- gapminder %>%
  select(country, continent) %>%
  rename(country_of_bean_origin = country)

gm_country <- gm_country %>%
  distinct(country_of_bean_origin, .keep_all = TRUE)

str(gm_country)
str(chocolate)

chocolate <- left_join(chocolate, gm_country, by = "country_of_bean_origin")

```


2.  Only keep reviews that have reviews from countries of bean origin with at least 10 reviews.
3.  Also, remove the country of bean origin named `"Blend"`.

**Hint**:

-   Check to see if there are any `NA`s in the new column. If there are any `NA`s, add the continent name for each row.

```{r}
choco_10reviews <-chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 10)

choco_10reviews %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    num_reviews = n(),
    unique_countries = n_distinct(country_of_bean_origin)
  )#50 countries left

#checking for missing data
sum(is.na(choco_10reviews$continent))
missing <- choco_10reviews %>% 
  filter(is.na(continent)) %>%
  select(country_of_bean_origin, continent)

missing %>%
  count(country_of_bean_origin)

unique(choco_10reviews$continent)

choco_10reviews <- choco_10reviews %>%
  mutate(continent = case_when(
    is.na(continent) & country_of_bean_origin %in% c("Belize", "Grenada", "St. Lucia", "Trinidad", "U.S.A.") ~ "Americas",
    is.na(continent) & country_of_bean_origin %in% c("Congo", "Sao Tome") ~ "Africa",
    is.na(continent) & country_of_bean_origin %in% c("Fiji", "Papua New Guinea", "Solomon Islands", "Vanuatu") ~ "Oceania",
    TRUE ~ continent)) %>% # Keep original continent if no match 
  filter(country_of_bean_origin != "Blend") 

(sum(is.na(choco_10reviews$continent))) #Now we have 0 missing values for continent
```

4.  Make a set of violin plots with ratings on the y-axis and `continent`s on the x-axis.

```{r}
ggplot(choco_10reviews) +
  aes(x = continent, y = rating) +
  geom_violin(adjust = 1.5)
```

# Part 3: Convert wide data into long data

The goal of this part of the assignment is to take a dataset that is either messy or simply not tidy and to make them tidy datasets. The objective is to gain some familiarity with the functions in the `dplyr`, `tidyr` packages. You may find it helpful to review the section on pivoting data from wide to long format and vice versa.

### Tasks

We are going to create a set of features for us to plot over time. Use the functions in `dplyr` and `tidyr` to perform the following steps to the `chocolate` dataset:

1.  Create a new set of columns titled `beans`, `sugar`, `cocoa_butter`, `vanilla`, `letchin`, and `salt` that contain a 1 or 0 representing whether or not that review for the chocolate bar contained that ingredient (1) or not (0).

```{r}

library(stringr)

choco_tidy <- chocolate %>%
  extract(
    col = ingredients,
    into = c("ingred_number", "ingred_list"),
    regex = "([1-5])-\\s*([A-Z,]+)"
  )

choco_tidy <- choco_tidy %>%
  mutate(
    beans = case_when(str_detect(ingred_list, "B") ~ 1, TRUE ~ 0),
    sugar = case_when(str_detect(ingred_list, "S") ~ 1, TRUE ~ 0),
    cocoa_butter = case_when(str_detect(ingred_list, "C") ~ 1, TRUE ~ 0),
    vanilla = case_when(str_detect(ingred_list, "V") ~ 1, TRUE ~ 0),
    lecithin = case_when(str_detect(ingred_list, "L") ~ 1, TRUE ~ 0),
    salt = case_when(str_detect(ingred_list, "Sa") ~ 1, TRUE ~ 0)  #There is no salt in the ingredients list
  )
```
2.  Create a new set of columns titled `char_cocoa`, `char_sweet`, `char_nutty`, `char_creamy`, `char_roasty`, `char_earthy` that contain a 1 or 0 representing whether or not that the most memorable characteristic for the chocolate bar had that word (1) or not (0). For example, if the word "sweet" appears in the `most_memorable_characteristics`, then record a 1, otherwise a 0 for that review in the `char_sweet` column (**Hint**: check out `str_detect()` from the `stringr` package).

```{r}
choco_tidy <- choco_tidy %>%
  mutate(
    char_cocoa = case_when(str_detect(most_memorable_characteristics, "cocoa") ~ 1, TRUE ~ 0),
    char_sweet = case_when(str_detect(most_memorable_characteristics, "sweet") ~ 1, TRUE ~ 0),
    char_nutty = case_when(str_detect(most_memorable_characteristics, "nutty") ~ 1, TRUE ~ 0),
    char_creamy = case_when(str_detect(most_memorable_characteristics, "creamy") ~ 1, TRUE ~ 0),
    char_roasty = case_when(str_detect(most_memorable_characteristics, "roasty") ~ 1, TRUE ~ 0),
    char_earthy = case_when(str_detect(most_memorable_characteristics, "earthy") ~ 1, TRUE ~ 0)
  )
```


3.  For each year (i.e. `review_date`), calculate the mean value in each new column you created across all reviews for that year. (**Hint**: If all has gone well thus far, you should have a dataset with 16 rows and 13 columns).

```{r}
ing_char <- choco_tidy %>%
  group_by(review_date) %>%
  summarize(beans = mean(beans),
            sugar = mean(sugar),
            cocoa_butter = mean(cocoa_butter),
            vanilla = mean(vanilla),
            lecithin = mean(lecithin),
            salt = mean(salt),
            char_cocoa = mean(char_cocoa),
            char_sweet = mean(char_sweet),
            char_nutty = mean(char_nutty),
            char_creamy = mean(char_creamy),
            char_roasty = mean(char_roasty),
            char_earthy = mean(char_earthy))

#A more concise option:

#ing_char <- choco_tidy %>%
#  group_by(review_date) %>%
#  summarize(across(c(beans, sugar, cocoa_butter, vanilla, lecithin, salt, 
#                     char_cocoa, char_sweet, char_nutty, char_creamy, 
#                     char_roasty, char_earthy), mean))
```

4.  Convert this wide dataset into a long dataset with a new `feature` and `mean_score` column.

It should look something like this:

```         
review_date     feature   mean_score
<dbl>           <chr>     <dbl>
2006    beans   0.967741935     
2006    sugar   0.967741935     
2006    cocoa_butter    0.903225806     
2006    vanilla 0.693548387     
2006    letchin 0.693548387     
2006    salt    0.000000000     
2006    char_cocoa  0.209677419     
2006    char_sweet  0.161290323     
2006    char_nutty  0.032258065     
2006    char_creamy 0.241935484 
```
### Notes

-   You may need to use functions outside these packages to obtain this result.

-   Do not worry about the ordering of the rows or columns. Depending on whether you use `gather()` or `pivot_longer()`, the order of your output may differ from what is printed above. As long as the result is a tidy data set, that is sufficient.

```{r}
ing_char <- ing_char %>%
  pivot_longer(cols = -review_date,
               names_to = "feature",
               values_to = "meanscore")
  
```
# Part 4: Data visualization

In this part of the project, we will continue to work with our now tidy song dataset from the previous part.

### Tasks

Use the functions in `ggplot2` package to make a scatter plot of the `mean_score`s (y-axis) over time (x-axis). There should be one plot (facet) for each `feature`. For full credit, your plot should include:

1.  An overall title for the plot and a subtitle summarizing key trends that you found. Also include a caption in the figure with your name.
2.  Both the observed points for the `mean_score`, but also a smoothed non-linear pattern of the trend
3.  All plots should be shown in the one figure
4.  There should be an informative x-axis and y-axis label

Consider playing around with the `theme()` function to make the figure shine, including playing with background colors, font, etc. See <https://github.com/lcolladotor/jhustatcomputing/issues/4> for an example on how you can choose colors if you have too many categories.

```{r}

ing_char$feature <- factor(ing_char$feature,
                           levels = c("char_cocoa", "char_creamy", "char_earthy", 
                                      "char_nutty", "char_roasty", "char_sweet", 
                                      "beans", "cocoa_butter", "lecithin", 
                                      "sugar", "vanilla", "salt"))

ggplot(ing_char, aes(x = review_date, y = meanscore)) +
  geom_point() +
  facet_wrap( ~ feature, ncol = 3) +
  geom_smooth() +
  ggtitle("Trends in Chocolate Characteristics and Ingredients") +
  theme(plot.title = element_text(hjust = 0.5))
```

